
---
title: "Analyzing the convergence of the SVNP_hyper" 
author: "Michael Koch"
date: "05-02-2025"
output: html_document
---


# Step1: Prepearations

```{r, message = F, warning = F}
# load packages outside of simulation
library(papaja)
#source required functions & parameters
getwd()
source('../../../R/functions.R')
source('../../../R/parameters.R')
```

```{r, cache = T}
# load results
resultsSVNP_hyper <- read_rds("../../../output/resultsSVNP_hyper.RDS")
convSVNP_hyper <- readr::read_rds("../../../output/convSVNP_hyper.RDS")
```

```{r}
# inspect a bit
nrow(resultsSVNP_hyper)
```

```{r}
(nrow(condPop)*nrow(condSVNP_hyper)*nIter) - nrow(resultsSVNP_hyper)
```

10 chains failed, which is in correspondence with the error R gave, when running the simulation on the server. Let 's find out the positions of those that failed. Perhaps they can be re-run. If they cannot, this may also be a result. 

ERROR message:

Error in checkForRemoteErrors(val) : 
  11 nodes produced errors; first error: subscript out of bounds
4.
stop(count, " nodes produced errors; first error: ", firstmsg,
domain = NA)
3.
checkForRemoteErrors(val)
2.
dynamicClusterApply(cl, fun, length(x), argfun)
1.
clusterApplyLB(clusters, 1:length(dataStanSVNP_hyper), sampling,
dataStan = dataStanSVNP_hyper, prior = "SVNP_hyper", modelPars = modelPars,
samplePars = samplePars)

```{r}
dataStanSVNP_hyper <- readr::read_rds("../../../data/dataStanSVNP_hyper.RDS")
n_cond <- length(dataStanSVNP_hyper)
```


```{r}
comp <- 1:n_cond
failedChainIndex <- which((comp %in% resultsSVNP_hyper$pos) == FALSE)
length(failedChainIndex)
failedChainIndex
```

```{r}
tibble(
  failedChainIndex = failedChainIndex,
  N = sapply(failedChainIndex, function(i) dataStanSVNP_hyper[[i]]$N),
  cross = sapply(failedChainIndex, function(i) dataStanSVNP_hyper[[i]]$cross)
)
```

The coondiitons under which it failed is almost always in 7/10 cases with N = 100 and cross = 0.2. once with N = 100 and cross = 0.5 and twice with N = 200 and cross = 0.2

# Convergence

Let's explore the convergence a bit (of the remaining set of conditions).

## Divergent transitions

```{r}
totalSumDiv <- sum(convSVNP_hyper$sumDiv)/2
totalSumDiv
totalSumDiv / (2000*nrow(convSVNP_hyper)/2)
```

In total, there are a 32851 divergent transition, but only a fraction of all transitions (0.0052) were divergent. Let's start by computing the percentage per individual replication. And then we can check how big that is on average per set of conditions.


```{r}
meanPropDiv <- 
convSVNP_hyper %>% 
  filter(parameter == "Rhat") %>% # doesn't matter which you pick here
  mutate(propDiv = sumDiv/ (samplePars$nSampling + samplePars$nWarmup)) %>% 
  group_by( N, cross) %>%
  summarise(meanPropDiv = mean(propDiv))
meanPropDiv
```

On average, they appear to be quite low. With N = 100 and cross = 0.2 the proportion is larger than for the rest, but still only at 0.5 % 

Let's find the indices of the runs where the divergent transitions exceeded 5%.

```{r}
indicesDIVHigher5Perc <- 
convSVNP_hyper %>% 
  filter(parameter == "Rhat") %>% # doesn't matter which you pick here
  mutate(propDIV = sumDiv/(samplePars$nSampling + samplePars$nWarmup)) %>% 
  filter(propDIV >= convCriteria$strict$maxPropDIV
) %>% 
  select(pos) %>% 
  pull()
length(indicesDIVHigher5Perc)
```

This would remove 19 runs:

```{r}
tibble(
  indicesDIVHigher5Perc = indicesDIVHigher5Perc,
  N = sapply(failedChainIndex, function(i) dataStanSVNP_hyper[[i]]$N),
  cross = sapply(failedChainIndex, function(i) dataStanSVNP_hyper[[i]]$cross)
)
```


## on average, this happened under no set of conditions.



```{r}
readr::write_rds(meanPropDIVlarger5, file = "~/1vs2StepBayesianRegSEM/Rmd/figures/meanPropDIVlarger5_SVNP_hyper.Rds")
```

## Effective Sample Size

Let's find the indices of all transitions where the effective sample size is less than 10% of the chain-length.

```{r}
indicesNEffTooSmall <- 
convSVNP_hyper %>% 
  filter(parameter == "n_eff") %>% 
  select(N, cross, pos, `lambdaMainC[1]`:`theta[6]`) %>% 
  mutate(across(-c(N, cross, pos), 
                ~ . / samplePars$nSampling
                )
         )%>% 
  filter_at(.vars = vars(`lambdaMainC[1]`:`theta[6]`), ~.x < convCriteria$strict$minPropNEFF) %>% 
  select(pos) %>% 
  pull()
```

```{r}
length(indicesNEffTooSmall)
```

By this you would lose 0 rows.


### R-Hat

Let's find the indices of all transitions where the R-hat is smaller than 1.05.

```{r}
indicesRHatTooSmall <- 
convSVNP_hyper %>% 
  filter(parameter == "Rhat") %>% 
  select(N, cross, pos, `lambdaMainC[1]`:`theta[6]`) %>% 
  # get indices of all iterations where for any parameter 
  filter_at(.vars = vars(`lambdaMainC[1]`:`theta[6]`), ~.x >= convCriteria$strict$maxRhat) %>% 
  select(pos) %>% 
  pull()
```

```{r}
length(indicesRHatTooSmall)
```

there are none.



